\graphicspath{ {images/} }

\titledquestion{Analyzing NMT Systems}[25]

\begin{parts}

    \part[3] Look at the {\monofam{src.vocab}} file for some examples of phrases and words in the source language vocabulary. When encoding an input Mandarin Chinese sequence into ``pieces'' in the vocabulary, the tokenizer maps the sequence to a series of vocabulary items, each consisting of one or more characters (thanks to the {\monofam{sentencepiece}} tokenizer, we can perform this segmentation even when the original text has no white space). Given this information, how could adding a 1D Convolutional layer after the embedding layer and before passing the embeddings into the bidirectional encoder help our NMT system? \textbf{Hint:} each Mandarin Chinese character is either an entire word or a morpheme in a word. Look up the meanings of 电, 脑, and 电脑 separately for an example. The characters 电 (electricity) and  脑 (brain) when combined into the phrase 电脑 mean computer.

    \ifans{As the embeddings are applied to "pieces" of the text, we may want to combine those pieces
    before inputting them into the bidirectional encoder. This would make the input to the bidirectional
    encoder potentially a word with the meaning intended from the context of the sentence. One way to 
    do this would be to use a Convolutional layer.}


    \part[8] Here we present a series of errors we found in the outputs of our NMT model (which is the same as the one you just trained). For each example of a reference (i.e., `gold') English translation, and NMT (i.e., `model') English translation, please:
    
    \begin{enumerate}
        \item Identify the error in the NMT translation.
        \item Provide possible reason(s) why the model may have made the error (either due to a specific linguistic construct or a specific model limitation).
        \item Describe one possible way we might alter the NMT system to fix the observed error. There are more than one possible fixes for an error. For example, it could be tweaking the size of the hidden layers or changing the attention mechanism.
    \end{enumerate}
    
    Below are the translations that you should analyze as described above. Only analyze the underlined error in each sentence. Rest assured that you don't need to know Mandarin to answer these questions. You just need to know English! If, however, you would like some additional color on the source sentences, feel free to use a resource like \url{https://www.archchinese.com/chinese_english_dictionary.html} to look up words. Feel free to search the training data file to have a better sense of how often certain characters occur.

    \begin{subparts}
        \subpart[2]
        \textbf{Source Sentence:} 贼人其后被警方拘捕及被判处盗窃罪名成立。 \newline
        \textbf{Reference Translation:} \textit{\underline{the culprits were} subsequently arrested and convicted.}\newline
        \textbf{NMT Translation:} \textit{\underline{the culprit was} subsequently arrested and sentenced to theft.}
        
        \ifans{
        \begin{enumerate}
            \item 'culprit' (singular) is used in the NMT translation instead of 'culprits' plural.
            \item Using the \href{https://www.archchinese.com/chinese_english_dictionary.html}{dictionary} 
        I couldn't find any words that indicated a plural for culprit. After a google search for how plurals
        work in Mandarin, I saw this \href{https://www.fluentu.com/blog/chinese/plural-in-chinese/}{article}.
        It says there is no direct analog of the 's' at the end of English words to
        indicate plural. Also, sometimes plurals are inferred from context. Therefore, I assume that the fact
        that there are many culprits is inferred from the context by the translator. The fact that the NMT system
        did not pick up this context may be due to how often 'culprit' (singular) occurs in the training set 
        versus 'culprits' (plural). Or it may be due to how often singular nouns appear vs plural nouns in
        the training set.
            \item To better use the context of the sentence to infer plurality, we could use a two-layer LSTM
        network. This may allow more complicated contextual references to be learned. 
        \end{enumerate}
        }

        \subpart[2]
        \textbf{Source Sentence}: 几乎已经没有地方容纳这些人,资源已经用尽。\newline
        \textbf{Reference Translation}: \textit{there is almost no space to accommodate these people, and resources have run out.   }\newline
        \textbf{NMT Translation}: \textit{the resources have been exhausted and \underline{resources have been exhausted}.}
        
        \ifans{}

        \subpart[2]
        \textbf{Source Sentence}: 当局已经宣布今天是国殇日。 \newline
        \textbf{Reference Translation}: \textit{authorities have announced \underline{a national mourning today.}}\newline
        \textbf{NMT Translation}: \textit{the administration has announced \underline{today's day.}}
        
        \ifans{}
        
        \subpart[2] 
        \textbf{Source Sentence\footnote{This is a Cantonese sentence! The data used in this assignment comes from GALE Phase 3, which is a compilation of news written in simplified Chinese from various sources scraped from the internet along with their translations. For more details, see \url{https://catalog.ldc.upenn.edu/LDC2017T02}. }:} 俗语有云:``唔做唔错"。\newline
        \textbf{Reference Translation:} \textit{\underline{`` act not, err not "}, so a saying goes.}\newline
        \textbf{NMT Translation:} \textit{as the saying goes, \underline{`` it's not wrong. "}}
        
        \ifans{}
    \end{subparts}


    \part[14] BLEU score is the most commonly used automatic evaluation metric for NMT systems. It is usually calculated across the entire test set, but here we will consider BLEU defined for a single example.\footnote{This definition of sentence-level BLEU score matches the \texttt{sentence\_bleu()} function in the \texttt{nltk} Python package. Note that the NLTK function is sensitive to capitalization. In this question, all text is lowercased, so capitalization is irrelevant. \\ \url{http://www.nltk.org/api/nltk.translate.html\#nltk.translate.bleu_score.sentence_bleu}
    } 
    Suppose we have a source sentence $\bs$, a set of $k$ reference translations $\br_1,\dots,\br_k$, and a candidate translation $\bc$. To compute the BLEU score of $\bc$, we first compute the \textit{modified $n$-gram precision} $p_n$ of $\bc$, for each of $n=1,2,3,4$, where $n$ is the $n$ in \href{https://en.wikipedia.org/wiki/N-gram}{n-gram}:
    \begin{align}
        p_n = \frac{ \displaystyle \sum_{\text{ngram} \in \bc} \min \bigg( \max_{i=1,\dots,k} \text{Count}_{\br_i}(\text{ngram}), \enspace \text{Count}_{\bc}(\text{ngram}) \bigg) }{\displaystyle \sum_{\text{ngram}\in \bc} \text{Count}_{\bc}(\text{ngram})}
    \end{align}
     Here, for each of the $n$-grams that appear in the candidate translation $\bc$, we count the maximum number of times it appears in any one reference translation, capped by the number of times it appears in $\bc$ (this is the numerator). We divide this by the number of $n$-grams in $\bc$ (denominator). \newline 

    Next, we compute the \textit{brevity penalty} BP. Let $len(c)$ be the length of $\bc$ and let $len(r)$ be the length of the reference translation that is closest to $len(c)$ (in the case of two equally-close reference translation lengths, choose $len(r)$ as the shorter one). 
    \begin{align}
        BP = 
        \begin{cases}
            1 & \text{if } len(c) \ge len(r) \\
            \exp \big( 1 - \frac{len(r)}{len(c)} \big) & \text{otherwise}
        \end{cases}
    \end{align}
    Lastly, the BLEU score for candidate $\bc$ with respect to $\br_1,\dots,\br_k$ is:
    \begin{align}
        BLEU = BP \times \exp \Big( \sum_{n=1}^4 \lambda_n \log p_n \Big)
    \end{align}
    where $\lambda_1,\lambda_2,\lambda_3,\lambda_4$ are weights that sum to 1. The $\log$ here is natural log.
    \newline
    \begin{subparts}
        \subpart[5] Please consider this example: \newline
        Source Sentence $\bs$: \textbf{需要有充足和可预测的资源。} 
        \newline
        Reference Translation $\br_1$: \textit{resources have to be sufficient and they have to be predictable}
        \newline
        Reference Translation $\br_2$: \textit{adequate and predictable resources are required}
        
        NMT Translation $\bc_1$: there is a need for adequate and predictable resources
        
        NMT Translation $\bc_2$: resources be suﬀicient and predictable to
        
        Please compute the BLEU scores for $\bc_1$ and $\bc_2$. Let $\lambda_i=0.5$ for $i\in\{1,2\}$ and $\lambda_i=0$ for $i\in\{3,4\}$ (\textbf{this means we ignore 3-grams and 4-grams}, i.e., don't compute $p_3$ or $p_4$). When computing BLEU scores, show your work (i.e., show your computed values for $p_1$, $p_2$, $len(c)$, $len(r)$ and $BP$). Note that the BLEU scores can be expressed between 0 and 1 or between 0 and 100. The code is using the 0 to 100 scale while in this question we are using the \textbf{0 to 1} scale. Please round your responses to 3 decimal places. 
        \newline
        
        Which of the two NMT translations is considered the better translation according to the BLEU Score? Do you agree that it is the better translation?
        
        \ifans{Let $f_c('n-gram') = \min(\max_{i=1,...,k}(\text{Count}_{r_i}(ngram),\text{Count}_c(ngram)))$.
        \fc1{is} = 1, \fc1{a} = 0, \fc1{need} = 0, \fc1{for} = 0, \fc1{adequate} = 1, \fc1{and} = 1, \fc1{predictable} = 1
        \fc1{resources} = 1, \fc1{there is} = 0,  \fc1{is a} = 0, \fc1{need for} = 0, \fc1{for adequate} = 0,
        \fc1{adequate and} = 1, \fc1{and predictable} = 1, \fc1{predicatable resources} = 1. So, for $c_1$, 
        $p_1 = 4/9$, $p_2 = 3/8$. $\text{len}(c_1)$ is closer to $\text{len}(r_1)$ than $\text{len}(r_2)$.

        \[\text{BLEU}_{c_1} = \exp\left(1 - \dfrac{11}{9}\right) \cdot \exp\left(\dfrac{1}{2}\log{\dfrac{4}{9}}+\dfrac{1}{2}\log\dfrac{3}{8}\right)
            = 0.45\]

        For $c_2$, \fc2{resources} = 1, \fc2{be} = 1, \fc2{sufficient} = 1, \fc2{and} = 1, \fc2{predictable}
        =1, \fc2{to} = 1, \fc2{resources be} = 0, \fc2{be sufficient} = 1, \fc2{sufficient and} = 1,
        \fc2{and predictable} = 1, \fc2{predictable to} = 0. $p_1 = 1$, $p_2 = 3/5$, $\text{len}(c_2) = \text{len}(r_2)$

        \[\text{BLEU}_{c_2} = \exp\left(\dfrac{1}{2}\log1 + \dfrac{1}{2}\log\dfrac{3}{5}\right) = 0.548\]

        $c_2$ is considered better by BLEU score. I do not agree that this is the better translation.
        }
        
        \subpart[5] Our hard drive was corrupted and we lost Reference Translation $\br_1$. Please recompute BLEU scores for $\bc_1$ and $\bc_2$, this time with respect to $\br_2$ only. Which of the two NMT translations now receives the higher BLEU score? Do you agree that it is the better translation?
        
        \ifans{Let $f_c('n-gram') = \min(\max_{i=1,...,k}(\text{Count}_{r_i}(ngram),\text{Count}_c(ngram)))$.
        \fc1{is} = 1, \fc1{a} = 0, \fc1{need} = 0, \fc1{for} = 0, \fc1{adequate} = 1, \fc1{and} = 1, \fc1{predictable} = 1
        \fc1{resources} = 1, \fc1{there is} = 0,  \fc1{is a} = 0, \fc1{need for} = 0, \fc1{for adequate} = 0,
        \fc1{adequate and} = 1, \fc1{and predictable} = 1, \fc1{predicatable resources} = 1. So, for $c_1$, 
        $p_1 = 4/9$, $p_2 = 3/8$.

        \[\text{BLEU}_{c_1} =  \exp\left(\dfrac{1}{2}\log{\dfrac{4}{9}}+\dfrac{1}{2}\log\dfrac{3}{8}\right)
            = 0.678\]

        For $c_2$, \fc2{resources} = 1, \fc2{be} = 0, \fc2{sufficient} = 0, \fc2{and} = 1, \fc2{predicatable} = 1,
        \fc2{to} = 0, \fc2{resources be} = 0, \fc2{be sufficient} = 0, \fc2{sufficient and} = 0, 
        \fc2{and predicatable} = 1, \fc2{predicatable to} = 0, $p_1 = 3/6$, $p_2 = 1/5$.

        \[\text{BLEU} = 1 \exp\left(\dfrac{1}{2}\log(\dfrac{3}{6}) + \dfrac{1}{2}\log(\dfrac{1}{5})\right) = 0.607\]

        $c_1$ is now better. I agree it is the better translation.
        }

        
        \subpart[2] Due to data availability, NMT systems are often evaluated with respect to only a single reference translation. Please explain (in a few sentences) why this may be problematic. In your explanation, discuss how the BLEU score metric assesses the quality of NMT translations when there are multiple reference transitions versus a single reference translation.
        
        \ifans{Different translations can use different words and phrases and still be good translations.
            When comparing candidate translations to multiple references using the BLEU score, each of the
            $p_i$, and hence the final BLEU score, can be high if words or phrases in the candidate are present
            in atleast one of the references. However, when comparing to a single reference, if the candidate
            translation uses different words or phrases from the reference which have an equivalent meaning, the candidate may receive a
            low BLEU score eventhough it may be a good translation.}
        
        \subpart[2] List two advantages and two disadvantages of BLEU, compared to human evaluation, as an evaluation metric for Machine Translation. 
        
        \ifans{Advantages:
        \begin{enumerate}
            \item Cheap to compute.
            \item Requires only reference translations to compute BLEU scores for a large number of candidates.
             Therefore, can be applied to a large number of candidates programmatically.
        \end{enumerate}
        Disadvantages:
        \begin{enumerate}
            \item As BLEU score compares words or phrases in candidate translations to references, it depends highly on the availability of
                high quality references. Such references may not always be available.
            \item Even with high quality references available, if the candidate uses words or phrases not present
                in any reference, it may receive a low BLEU score even if its meaning is equivalent to the reference
                translations.
        \end{enumerate}
        }
    \end{subparts}
\end{parts}
